---
title: "Examples"
author: Marco Pittarello
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(iPastoralist)


```

```{r include=FALSE}
library(knitr)
library(kableExtra)
library(dendextend)
library(iPastoralist)
library(amap)

load(file = "data.RData")
data<-as.data.frame(data)
vegetation<-data[,c(2,7:30)]


```

# Example 1 - use of 'vegetation_abundance' function

Suppose we want to convert Frequency of occurrence (FO) to Species percentage cover (%SC), considering also occasional species. As the %SC for each survey will (likely) be greater than 100, we want to rescale %SC of each species per each survey to obtain a sum of 100 (i.e. a proportion of %SC). The dataset used as example is shown in "Data input format" vignette, where the total measurements per transect is 25. Therefore, to obtain \$SC FO should be multiplied by **4** (so that they refer to 100 measurements.)

```{r}
library(iPastoralist)
vegetation.sc<-vegetation_abundance(database = vegetation,
                                    species.cover.coefficient = 4,
                                    method = "SRA_SC.fo.occ")
```

    head(vegetation.sc)

```{r echo=FALSE}
kable(head(vegetation.sc[,1:15]),align = 'c')%>%
  kable_styling(font_size = 12)
```

we can check that the sum of %SC for each survey is 100

```{r}
colSums(vegetation.sc)
```

# Example 2 - computation of ecological indexes

In this case we want to compute the average Landolt indicator values for each survey, weighted with species abundance.

If the occasional species are not considered, the SRA will be used. Conversely, if we would like to keep into account also occasional species, the SRA will be calculated with the %SC rescaled to 100 (more detail in "vegetation_abundance" function).

In this case we will consider also occasional species.

The input database is the one with the Frequency of occurrences, i.e. the dataframe used in this tutorial named "vegetation".

```{r}
ec.index<-ecological_indexes(database.vegetation = vegetation,
                             database.indexes = data[,c("F_Landolt","R_Landolt","N_Landolt")],
                             occasional.species = TRUE,
                             species.cover.coefficient = 4,
                             weight = TRUE)
```

Notes about the "ecological_indexes" function:

-   **database.indexes** = database with Ecological indicators, without the column of species names. NA values must indicated as 999
-   **occasional.species** = Logical. TRUE if you want to take into account occasional species.
-   **species.cover.coefficient** = only if "occasional.species=TRUE". Coefficient that multiplies FO so that the number of total touches refer to 100
-   **weight**: Logical. TRUE if you want to weight Ecological indicators with abundance.

the output will be as follow:

```{r}
ec.index
```

# Example 3 - Matching a dendrogram with vegetation data

The aim of this example is to generate a cluster analysis of vegetation data, with associated to the dendrogram the species of each survey ordered by their abundance. This approach would help where to cut the dendrogram.

From the dataset shown in 'Data input format' vignette, the columns related to plant species names and all surveys are selected.

```{r echo=TRUE}
vegetation<-data[,c(2,7:30)]

```

```{r echo=FALSE}
kable(head(vegetation[,1:15]),align = 'c')%>%
  kable_styling(font_size = 12)
```

Then, we need to compute the Species Relative Abundance

```{r}
sra<-vegetation_abundance(database = vegetation,method = "SRA_fo")
```

    head(sra)

```{r echo=FALSE}
kable(head(sra[,1:15]),align = 'c')%>%
  kable_styling(font_size = 12)
```

Now we can generate a dendrogram using the **hclust** function of the R 'stats' base package.

```{r fig.height=4, fig.width=7}
db.dendro<-t(sra)#for the cluster analysis the database has to be transposed so to have surveys on rows and species on columns

library(amap)#package for computing a distance matrix
d <- Dist(db.dendro, method = "correlation")#distance matrix
cluster <- hclust(d, method="average") # clustering method

#plotting the dendrogram
par(cex=0.5, mar=c(5, 8, 4, 1))#set label size
plot(cluster,cex=0.8,hang =-1,
     main=paste0("dMatrix: Correlation","---","clustMet: UPGMA"))

```

To extract the first ten species ordered decreasingly by their SRA for each survey, we can run the **'first_ten_species'** function:

```{r}
firstTenSpecies<-first_ten_species(data_SRA_SC = sra ,#database with species abundance. In this example is the 'sra' database 
                                   join.dendrogram = TRUE,#LOGICAL. TRUE if species abundance need to be joined with a dendrogram
                                   cluster.hclust = cluster)# object of containing the output from 'hclust' function of 'stats' package


```

```{r echo=FALSE}
kable(head(firstTenSpecies[,1:10],15),align='c')%>%
  kable_styling(font_size = 12)
```

This output can be graphically merged with the dendrogram. First of all it is needed to create a database for running the cluster analysis with (e.g. four) species ordered by abundance as row labels, with the use of `dendspe` function:

```{r}
db.dendroLab<-denspe(nspe=4,
                     SpeAbund = firstTenSpecies,
                     dbClust = db.dendro)
```

Then, the cluster analysis can be run again by using the new database and then plot the dendrogram:

```{r fig.height=7, fig.width=7}
d.lab <- Dist(db.dendroLab,#database name
           method= "correlation") 
cluster.lab <- hclust(d.lab,#distance matrix
                   method="average") 

cluster.plot<-as.dendrogram(cluster.lab)#coversion of cluster to class 'dendrogram' of 'dendexted' package (which allows a better customisation of the plot)

par(mfrow=c(1,1),mar=c(1,2,2,18))#graphical settings
plot(cluster.plot %>% 
  set("labels_cex", 0.6),horiz=T)
```

From this view, it can be easier to identify survey groups based on their vegetation composition and abundance. Suppose here to identify 8 groups, that can be visually shown with the following base plotting functions:

```{r fig.height=7, fig.width=7}
par(mfrow=c(1,1),mar=c(1,2,2,18))#graphical settings
cluster.plot %>% 
  set("labels_cex", 0.6) %>% 
  set("branches_k_color", 
      k = 8) %>% 
  plot(main = "Dendrogram",horiz=T)
```

Now, for each group it would be helpful to compute the average composition. First of all we can extract the ID group for each survey with the **'clustOrder'** function:

```{r}
group.id<-clustOrder(cluster.hclust = cluster,# object of containing the output from 'hclust' function of 'stats' package
                      cluster.group = T, #Logical. TRUE if you want to specify the number of groups of the dendrogram. 
                      cluster.number = 8) # Specify the number of groups. 
```

```{r echo=FALSE}
kable(head(group.id,10),align='c')%>%
  kable_styling(font_size = 12)
```

The 'cluster' column is the one according to which the vegetation surveys will be pooled together. Therefore, we first need to add this column to the database with the abundances used as input of the dendrogram (i.e. 'db.dendro' dataset)

```{r}
dendro.merge<-merge(group.id,db.dendro,by.x = "Survey",by.y = "row.names")
row.names(dendro.merge)<-dendro.merge[,1]#renaming dataset rows
dendro.merge1<-dendro.merge[,-c(1,3)]#deleting useless columns
```

```{r echo=FALSE}
kable(head(dendro.merge1[,1:15],10),align='c')%>%
  kable_styling(font_size = 12)
```

Lastly, with the **'clustGroupAggregate'** function, the average composition for each group can be computed:

```{r}
aggregate<-clustGroupAggregate(dendro.merge1)
```

```{r echo=FALSE}
kable(head(aggregate[,1:15],10),align='c')%>%
  kable_styling(font_size = 12)
```
